A VLT/FORS2 Narrowband Imaging Search for \ion{Mg}{2} Emission Around $\lowercase{z}\sim0.7$ Galaxies

### 2018-09-09 for_submission/rvaught_2018.tex
update 9/2/18
Sept 8th

\begin{document}
\title{A VLT/FORS2 Narrowband Imaging Search for \ion{Mg}{2} Emission Around $\lowercase{z}\sim0.7$ Galaxies }
\author{Ryan Rickards Vaught}
\affiliation{Department of Astronomy, San Diego State University, San Diego, CA 92182}
\affiliation{Department of Physics, University of California, San Diego, 9500 Gilman Dr., La Jolla, CA 92093, USA}

 \author{Kate H. R. Rubin }
 \affiliation{Department of Astronomy, San Diego State University, San Diego, CA 92182, USA}

@@

 \author{J. Xavier Prochaska}
 \affiliation{Department of Astronomy \& Astrophysics, University of California, 1156 High St., Santa Cruz, CA 95064 USA}

\author{Joseph F. Hennawi}
\affiliation{Department of Physics, Broida Hall, University of California, Santa Barbara, CA 93106-9530}

@@

%\begin{table*}[t]
%\centering
%\caption{Properties of the 5 galaxies in our sample as estimated in \cite{Rubin_2014}. The EW (in the observed frame) includes both components of the \ion{Mg}{2} doublet and is 
%determined from analysis of the supplemental Keck/LRIS spectra.}
%\begin{tabular}{llllllll} \hline \hline
%Object\footnote{ Name in parenthesis to be used as shorthand for the objects throughout the remainder of the paper. } & R.A. & Dec  & $z$ & SFR($M_{\odot}$ yr$^{-1}$) & $\log{M_{*}/M_{\odot}}$ & EW$_{\rm{obs}}$(\AA) & $\tau_V$\footnote{ Total V-band optical depth of dust as estimated by SED fitting with MAGPHYS. }\smallskip      \\ \hline 
%J033225.26-274524.0 (J.26)      & 03:32:25.26 & -27:45:23.9 & 0.6660 & $9.1_{-3.7}^{+1.3}$& $9.86_{-0.04}^{+0.05}$ & $7.539\pm 0.354$ &  $1.227_{-0.20}^{+1.54}$\\

%J033231.36-274725.0 (J.36)      & 03:32:31.35 & -27:47:24.9 &   0.6669 & $10.5_{-1.6}^{+1.7}$ & $10.02_{-0.03}^{+0.03}$&$5.835 \pm 0.493$ & $1.377_{-0.23}^{+0.60}$\\

%J033230.03-274347.3  (J.03)     & 03:32:30.03 & -27:43:47.2  &   0.6679 & $3.8_{-0}^{+0}$ & $10.98_{-0.0}^{+0.01}$ &$12.794 \pm 1.710$ & $0.297_{-0.0}^{+0.0}$ \\

%J033229.64-274242.6  (J.64)    & 03:32:29.64 & -27:42:42.5 & 0.6671 & $40.5_{-12.1}^{+8.2}$ & $10.30_{-0.03}^{+0.07}$ &$13.239 \pm 0.263$ & $3.897_{-0.93}^{+1.15}$\\

%J033230.57-274518.2  (J.57)    & 03:32:30.56 & -27:45:18.2 &   0.6807  & $12.6_{-2.1}^{+1.7}$ & $10.48_{-0.07}^{+0.03}$ &$6.106 \pm 0.370$ & $1.262_{-0.40}^{+1.23}$ \\
%\hline 
%\end{tabular}
%\tablenotemark{}
%\tablenotetext{}{\textit{Note} - The uncertainty intervals reported for SFR, $\log M_*/M_{\odot}$, and $\tau_V$ values indicate the $\pm34$th-percentile values of the posterior probability distribution function of each quantity as determined via modeling of the objects' SEDs with MAGPHYS (\citeauthor{daCunha2008} \citeyear{daCunha2008}, \citeyear{daCunha2012}) as described in \citet{Rubin_2014}}
%\label{tab:prop}
%\end{table*}

\begin{deluxetable*}{c c c c D D c D}
\tablecaption{Properties of the 5 galaxies in our sample as estimated in \cite{Rubin_2014}. The EW (in the observed frame) includes both components of the \ion{Mg}{2} doublet and is 
determined from analysis of the supplemental Keck/LRIS spectra. \label{tab:prop}}
\tablehead{\colhead{Object} & \colhead{R.A.} & \colhead{Dec} & \colhead{$z$} & \multicolumn2c{SFR} & \multicolumn2c{$\log{M}$} & \colhead{EW$_{\rm{obs}}$} & \multicolumn2c{ $\tau_V$} \\ 
\colhead {} & \colhead {} & \colhead {} & \colhead{} & \multicolumn2c{($M_{\odot}$ yr$^{-1}$)} & \multicolumn2c{$\log{M_{*}/M_{\odot}}$} & \colhead{(\AA)} & \multicolumn2c{}}
\decimals
\startdata
J033225.26-274524.0 (J.26)     & 03:32:25.26 & -27:45:23.9  &   0.6660  & $9.1_{-3.7}^{+1.3}$     & $9.86_{-0.04}^{+0.05}$   & \ $7.539\pm 0.354$   &  $1.227_{-0.20}^{+1.54}$ \\ 
J033231.36-274725.0 (J.36)     & 03:32:31.35 & -27:47:24.9  &   0.6669  & $10.5_{-1.6}^{+1.7}$   & $10.02_{-0.03}^{+0.03}$ & \ $5.835 \pm 0.493$   & $1.377_{-0.23}^{+0.60}$ \\
J033230.03-274347.3 (J.03)     & 03:32:30.03 & -27:43:47.2  &   0.6679  & $3.8_{-0}^{+0}$           & $10.98_{-0.0}^{+0.01}$   & $12.794 \pm 1.710$ & $0.297_{-0.0}^{+0.0}$ \\
J033229.64-274242.6 (J.64)     & 03:32:29.64 & -27:42:42.5  &   0.6671  & $40.5_{-12.1}^{+8.2}$ & $10.30_{-0.03}^{+0.07}$ & $13.239 \pm 0.263$ & $3.897_{-0.93}^{+1.15}$ \\
J033230.57-274518.2 (J.57)     & 03:32:30.56 & -27:45:18.2  &   0.6807  & $12.6_{-2.1}^{+1.7}$   & $10.48_{-0.07}^{+0.03}$ & \ $6.106 \pm 0.370$   & $1.262_{-0.40}^{+1.23}$ \\
\enddata
\tablecomments{The uncertainty intervals reported for SFR, $\log M_*/M_{\odot}$, and $\tau_V$ values indicate the $\pm34$th-percentile values of the posterior probability distribution function of each quantity as determined via modeling of the objects' SEDs with MAGPHYS (\citeauthor{daCunha2008} \citeyear{daCunha2008}, \citeyear{daCunha2012}) as described in \citet{Rubin_2014}}
\end{deluxetable*}

@@

\subsection{Absolute Flux Calibration}
We acquired observations of the standard star GD50 from archival ESO calibration imaging at 4 independent epochs. Performing aperture photometry at each epoch and airmass, we calculated the atmospheric extinction coefficients, $k$, to be 0.181 magnitudes for the HeII/3000+48 filter and 0.190 magnitudes for the HeII+47 filter. We perform absolute flux calibration using the methods of \cite{Jacoby1987}. We first convolve the spectral energy distribution of the standard star, $F(\lambda)$ in ergs sec$^{-1}$ \AA$^{-1}$ cm$^{-2}$, with that of the known transmission curve of the filter, $T_{i}(\lambda)$. This yields $F_i$, the total observable flux in each bandpass filter $i$ with units of ergs sec$^{-1}$ cm$^{-2}$:
\begin{equation*}
F_{i}=\int F(\lambda)T_{i}(\lambda)d\lambda.
\end{equation*}
It is not uncommon to assume that $F(\lambda)$ is constant over the small width of the filter. 
However, since our filter transmission curves are sampled at wavelengths similar to sampling of the spectrum of GD50 from the the CALSPEC library \citep{Bohlin2017}, we interpolate both spectra and compute the integral numerically. 
The conversion from count rate to flux units for each filter is then given by
\begin{equation*}
S_{i}=\dfrac{F_{i}}{C10^{k_{i}A}},
\end{equation*}\\
where $k_i$ is the extinction in magnitudes per airmass, A is the airmass for each individual exposure, C is the measured count rate of the standard star and $S_i$ is in units of ergs counts$^{-1}$ cm$^{-2}$. Before image co-addition, each science image is corrected for atmospheric extinction by multiplying each frame by $10^{k_{i}A}$. Next, the image is divided by the exposure time, effectively putting the image in units of counts per sec. After co-addition, the images are then multiplied by the appropriate sensitivity factor $S_{i}$. This puts the final images in the appropriate flux units, ergs sec $^{-1}$ cm$^{-2}$.

@@

In the case of perfect sky subtraction and continuum subtraction, the 1$\sigma$ SB limit for an extended source is $SB_{1}/\sqrt{A_\text{src}}$, where $A_\text{src}$ is the area in arcsec$^2$ and $SB_{1}$ is the surface brightness limit per 1 arcsec$^2$ aperture. However, our actual detection limits are altered by systematic error from imperfect subtraction. Therefore, we determine the limits as follows. We first mask all the artifacts and sources in the continuum-subtracted images. Next, we generate many apertures with sizes similar to our extended annulus ($\sim 20 \rm\ sq.arcsec$), place them at random, and extract the fluxes, $F_{\text{src}}$, within these apertures. We then normalize the vales of $F_{\text{src}}$ by dividing by $\sigma_{\text{src}}$, where $\sigma_{\text{src}} \equiv SB_{1}\sqrt{A_\text{src}}$. For perfect sky subtraction and continuum subtraction, the distribution of extracted fluxes should follow a Gaussian distribution with a variance equal to $\sigma_{\text{src}}$. The distribution of $F_{\text{src}}/\sigma_{\text{src}}$ for these apertures is shown in Figure \ref{fig:limits}

@@

We construct smoothed $\chi$ images, $\chi_{\text{smooth}}$, for each level of simulated emission, following the techniques in \cite{Hennawi2013} and \cite{Battaia_2015}. These $\chi_{\text{smooth}}$ images can aid in identifying the presence and detectability of extended \ion{Mg}{2} emission.

First, the model continuum-subtracted image is smoothed by the operation:
\begin{equation}
I_{\text{smooth}}= \text{CONVOLVE[line-continuum]},
\end{equation}
where the CONVOLVE operation indicates convolution of the \ion{Mg}{2} images with a Gaussian kernel with FWHM=1.5 pixels. Next, we computed the sigma image, $\sigma_{\text{smooth}}$, by convolving the propagated error image:
\begin{equation}
\sigma_{\text{smooth}}=\sqrt{\text{CONVOLVE}^2[\sigma^2]},
\end{equation}
where the CONVOLVE$^2$ operation indicates convolution of the image with the square of the Gaussian kernel. The smoothed $\chi$ image is then the product of the data image divided by the sigma image.
\begin{equation}
\chi_{\text{smooth}}=\frac{I_{\text{smooth}}}{\sigma_{\text{smooth}}}.
\end{equation}

>>>>

\begin{document}
\title{A VLT/FORS2 Narrowband Imaging Search for \ion{Mg}{2} Emission Around $\lowercase{z}\sim0.7$ Galaxies }
\author{Ryan Rickards Vaught}
\affiliation{Department of Astronomy, San Diego State University, San Diego, CA 92182, USA}
\affiliation{Department of Physics, University of California, San Diego, 9500 Gilman Dr., La Jolla, CA 92093, USA}

 \author{Kate H. R. Rubin}
 \affiliation{Department of Astronomy, San Diego State University, San Diego, CA 92182, USA}

@@

 \author{J. Xavier Prochaska}
 \affiliation{Department of Astronomy \& Astrophysics, University of California, 1156 High Street, Santa Cruz, CA 95064, USA}

\author{Joseph F. Hennawi}
\affiliation{Department of Physics, Broida Hall, University of California, Santa Barbara, CA 93106, USA}

@@

\begin{deluxetable*}{c c c c D D c D}
\tablecaption{Properties of the 5 galaxies in our sample as estimated in \cite{Rubin_2014}. The EW (in the observed frame) includes both components of the \ion{Mg}{2} doublet and is 
determined from analysis of the supplemental Keck/LRIS spectra. \label{tab:prop}}
\tablehead{\colhead{Object} & \colhead{R.A.} & \colhead{Dec} & \colhead{$z$} & \multicolumn2c{$\log$(SFR)} & \multicolumn2c{$\log{(M)}$} & \colhead{EW$_{\rm{obs}}$} & \multicolumn2c{ $\tau_V$} \\ 
\colhead {} & \colhead {} & \colhead {} & \colhead{} & \multicolumn2c{($\log M_{\odot}$ yr$^{-1}$)} & \multicolumn2c{($\log{M_{*}/M_{\odot}}$)} & \colhead{(\AA)} & \multicolumn2c{}}
\decimals
\startdata
J033225.26-274524.0 (J.26)     & 03:32:25.26 & -27:45:23.9  &   0.6660  & $9.1_{-3.7}^{+1.3}$     & $9.86_{-0.04}^{+0.05}$   & \ $7.539\pm 0.354$   &  $1.227_{-0.20}^{+1.54}$ \\ 
J033231.36-274725.0 (J.36)     & 03:32:31.35 & -27:47:24.9  &   0.6669  & $10.5_{-1.6}^{+1.7}$   & $10.02_{-0.03}^{+0.03}$ & \ $5.835 \pm 0.493$   & $1.377_{-0.23}^{+0.60}$ \\
J033230.03-274347.3 (J.03)     & 03:32:30.03 & -27:43:47.2  &   0.6679  & $3.8_{-0}^{+0}$           & $10.98_{-0.0}^{+0.01}$   & $12.794 \pm 1.710$ & $0.297_{-0.0}^{+0.0}$ \\
J033229.64-274242.6 (J.64)     & 03:32:29.64 & -27:42:42.5  &   0.6671  & $40.5_{-12.1}^{+8.2}$ & $10.30_{-0.03}^{+0.07}$ & $13.239 \pm 0.263$ & $3.897_{-0.93}^{+1.15}$ \\
J033230.57-274518.2 (J.57)     & 03:32:30.56 & -27:45:18.2  &   0.6807  & $12.6_{-2.1}^{+1.7}$   & $10.48_{-0.07}^{+0.03}$ & \ $6.106 \pm 0.370$   & $1.262_{-0.40}^{+1.23}$ \\
\enddata
\tablecomments{The uncertainty intervals reported for SFR, $\log M_*/M_{\odot}$, and $\tau_V$ values indicate the $\pm34$th-percentile values of the posterior probability distribution function of each quantity as determined via modeling of the objects' SEDs with MAGPHYS (\citeauthor{daCunha2008} \citeyear{daCunha2008}, \citeyear{daCunha2012}) as described in \citet{Rubin_2014}}
\end{deluxetable*}

@@

\subsection{Absolute Flux Calibration}
We acquired observations of the standard star GD50 from archival ESO calibration imaging at 4 independent epochs. Performing aperture photometry at each epoch and airmass, we calculated the atmospheric extinction coefficients, $k$, to be 0.181 magnitudes for the HeII/3000+48 filter and 0.190 magnitudes for the HeII+47 filter. We perform absolute flux calibration using the methods of \cite{Jacoby1987}. We first convolve the spectral energy distribution of the standard star, $F(\lambda)$ in ergs sec$^{-1}$ \AA$^{-1}$ cm$^{-2}$, with that of the known transmission curve of the filter, $T_{i}(\lambda)$. This yields $F_i$, the total observable flux in each bandpass filter $i$ with units of ergs sec$^{-1}$ cm$^{-2}$:
\begin{equation*}
F_{i}=\int F(\lambda)T_{i}(\lambda)d\lambda.
\end{equation*}
It is not uncommon to assume that $F(\lambda)$ is constant over the small width of the filter. 
However, since our filter transmission curves are sampled at wavelengths similar to sampling of the spectrum of GD50 from the the CALSPEC library \citep{Bohlin2017}, we interpolate both spectra and compute the integral without the constant $F(\lambda)$ assumption . 
The conversion from count rate to flux units for each filter is then given by
\begin{equation*}
S_{i}=\dfrac{F_{i}}{C10^{k_{i}A}},
\end{equation*}\\
where $k_i$ is the extinction in magnitudes per airmass, A is the airmass for each individual exposure, C is the measured count rate of the standard star and $S_i$ is in units of ergs counts$^{-1}$ cm$^{-2}$. Before image co-addition, each science image is corrected for atmospheric extinction by multiplying each frame by $10^{k_{i}A}$. Next, the image is divided by the exposure time, effectively putting the image in units of counts per sec. After co-addition, the images are then multiplied by the appropriate sensitivity factor $S_{i}$. This puts the final images in the appropriate flux units, ergs sec $^{-1}$ cm$^{-2}$.

@@

In the case of perfect sky subtraction and continuum subtraction, the 1$\sigma$ SB limit for an extended source is $SB_{1}/\sqrt{A_\text{src}}$, where $A_\text{src}$ is the area in arcsec$^2$ and $SB_{1}$ is the surface brightness limit per 1 arcsec$^2$ aperture. However, our actual detection limits are altered by systematic error from imperfect subtraction. Therefore, we determine the limits as follows. 
We first mask all the artifacts and sources in the continuum-subtracted images. Next, we generate many apertures with sizes similar to our extended annulus ($\sim 20 \rm\ sq.arcsec$), place them at random, and extract the fluxes, $F_{\text{src}}$, within these apertures. We then normalize the values of $F_{\text{src}}$ by dividing by $\sigma_{\text{src}}$, where $\sigma_{\text{src}} \equiv SB_{1}\sqrt{A_\text{src}}$. For perfect sky subtraction and continuum subtraction, the distribution of extracted fluxes should follow a Gaussian distribution with a variance equal to $\sigma_{\text{src}}$. The distribution of $F_{\text{src}}/\sigma_{\text{src}}$ for these apertures is shown in Figure \ref{fig:limits}

@@

To aid in identifying the presence and detectability of extended \ion{Mg}{2} emission, we construct a $\chi_{\text{smooth}}$, for each level of simulated emission, following the techniques in \cite{Hennawi2013} and \cite{Battaia_2015}.

To construct this smoothed image, we first performed the following operation on a continuum-subtracted image:
\begin{equation}
I_{\text{smooth}}= \text{CONVOLVE[line-continuum]},
\end{equation}
where the CONVOLVE operation indicates convolution of the \ion{Mg}{2} images with a Gaussian kernel with FWHM=1.5 pixels. Next, we computed the sigma image, $\sigma_{\text{smooth}}$, by convolving the propagated error image:
\begin{equation}
\sigma_{\text{smooth}}=\sqrt{\text{CONVOLVE}^2[\sigma^2]},
\end{equation}
where the CONVOLVE$^2$ operation indicates convolution of the image with the square of the Gaussian kernel. The smoothed $\chi$ image is then the product of the data image divided by the sigma image.
\begin{equation}
\chi_{\text{smooth}}=\frac{I_{\text{smooth}}}{\sigma_{\text{smooth}}}.
\end{equation}
